---
title: "Longer vs Shorter"
author: "Ziwei Zhao"
date: "May 6, 2024"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r echo=FALSE, warning=FALSE}
library(tidyverse)
library(survival)
library(ggplot2)
library(haven)
library(data.table)
```


```{r}
# Baseline data obtained using codes in STATA ClearGSK3_csv_BL.do file
data.BL <- read_dta("./GSK3_BL.dta") # Specific documentation is in "Description of GSK3 data.docx"

# Simplify, rename, and take information necessary for this research
data.BL2 <- data.BL%>%
  mutate(d1 = DATE_DOSE1, # date of dose1
         d2 = DATE_DOSE2, 
         d3 = DATE_DOSE3,
         d4 = DATE_DOSE4,
         DOE = STRTITT, # start date of follow-up period for cohort ITT (days since 1960-01-01)
         end_dof = ENDITTSE, # end date of follow-up period for cohort ITT (days since 1960-01-01)
         end_ef = ENDITTEE, # end date of extended follow-up period for cohort ITT (days since 1960-01-01)
         gender = SEX)%>%
  dplyr::select(PID, GROUP_NB, gender, AGEC, CENTER, d1:end_ef)


#Filter per-protocol polulation
#R3R -- 1
#R3C -- 2
#C3C -- 3
pid <- read.csv("./gsk_110021_pid_v02.csv", sep = ",")
table(pid$CENTER)
table(pid$SITEC2)
table(pid$CTY_NAM)
Tanzania <- filter(pid, CTY_NAM == "Tanzania")
table(Tanzania$CENTER)
```


Center:  
2   Kenya
3   Lambarene, Gabon
7   Manghica, Mozambique
8   Ghana
9   Lilongwe, Malawi
10  Tanzania
11  Nanoro, Burkina Faso
13  Kenya
15  Kenya
16  Tanzania
19  Ghana


PER PROTOCOL POPULATION

```{r}
# Include only subjects received first 3 doses, remove Manghica due to seasonal variation
children.c <- filter(data.BL2, AGEC == 1 & GROUP_NB ==3)%>% #2974
   filter(is.na(d3)== F & is.na(d2)== F & is.na(d1)== F)%>%
   filter(CENTER!=7)%>% #2546
   mutate(d1_m = as.double(difftime(as.Date(d1), as.Date("1960-01-01"), units = "days"))*12/365.25,
         d2_m = as.double(difftime(as.Date(d2), as.Date("1960-01-01"), units = "days"))*12/365.25,
         d3_m =  as.double(difftime(as.Date(d3), as.Date("1960-01-01"), units = "days"))*12/365.25,
         d4_m =  as.double(difftime(as.Date(d4), as.Date("1960-01-01"), units = "days"))*12/365.25,
         d1d2 = d2_m - d1_m,
         d2d3 = d3_m - d2_m,
         d3d4 = d4_m - d3_m,
         d3_td = as.double(difftime(as.Date(d3), as.Date("1960-01-01"), units = "days"))) %>%
  # remove out of interval doses
  filter(d1d2 > 0.6 & d1d2 < 1.4 & d2d3 > 0.6 & d2d3 < 1.4 & end_dof > d3_td + 14) #2344

children.i <- filter(data.BL2, AGEC == 1 & GROUP_NB !=3)%>% #5948
   filter(is.na(d3)== F & is.na(d2)== F & is.na(d1)== F)%>%
   filter(CENTER!=7)%>% #5027
   mutate(d1_m = as.double(difftime(as.Date(d1), as.Date("1960-01-01"), units = "days"))*12/365.25,
         d2_m = as.double(difftime(as.Date(d2), as.Date("1960-01-01"), units = "days"))*12/365.25,
         d3_m =  as.double(difftime(as.Date(d3), as.Date("1960-01-01"), units = "days"))*12/365.25,
         d4_m =  as.double(difftime(as.Date(d4), as.Date("1960-01-01"), units = "days"))*12/365.25,
         d1d2 = d2_m - d1_m,
         d2d3 = d3_m - d2_m,
         d3d4 = d4_m - d3_m,
         d3_td = as.double(difftime(as.Date(d3), as.Date("1960-01-01"), units = "days"))) %>%
  # remove out of interval doses
  filter(d1d2 > 0.6 & d1d2 < 1.4 & d2d3 > 0.6 & d2d3 < 1.4 & end_dof > d3_td + 14) #4593

data.BL.PP <- rbind(children.c, children.i)%>%
  arrange(PID)
```




Look at malaria events
```{r}
event.clinical <- read.csv("./gsk_110021_efficacy_v02.csv", sep = ",")%>%
  filter(TEMP >= 37.5 & FINALDEN > 5000)%>%
  mutate(date = strptime(DATE, format = "%d%b%Y"))%>%
  dplyr::select(PID, date, everything())

event.severe <- read.csv("./gsk_110021_sevmal_v02.csv", sep = ",")%>%
  mutate(date = strptime(DATE, format = "%d%b%Y"))%>%
  dplyr::select(PID, date, everything())
  
```



```{r}
# Write a function to truncate the data to certain length of follow-up time
function.finaldata <- function (month) {
  data.join <- left_join(x = data.BL.PP, y = event.clinical, by = "PID")%>%
   arrange(PID, date)%>%
   mutate(event = case_when(is.na(date)==T ~ 0,
                           TRUE ~ 1),
           d3_td = case_when(is.na(d3) == F ~ difftime(as.Date(d3), as.Date("1960-01-01"), units = "days"),
                       TRUE ~ max(
                         difftime(as.Date(d1), as.Date("1960-01-01"), units = "days"),
                         difftime(as.Date(d2), as.Date("1960-01-01"), units = "days"))),
          DOE = d3_td +14,
          event_td = difftime(as.Date(date), as.Date("1960-01-01"), units = "days"))%>%
   group_by(PID)%>% 
   distinct(event_td, .keep_all = T)%>% # emove duplicated event date for each subject
   ungroup()%>%
   mutate(
    # *time* is either event time or follow-up time if there's no events
         time = case_when(event != 0 ~ (event_td - (d3_td +14))*12/365.25,
                          TRUE ~ (end_dof - (d3_td+14))*12/365.25 ))
  #get pid for subjects who only have one event
  pid.one <- data.join%>%
    group_by(PID)%>%filter(n()==1)
data.join <- data.join%>%
  # deal with subjects experience only 1 event before DOE
  filter(time > 0 | PID %in% pid.one$PID)%>%
   mutate(event = case_when(time <= 0 ~ 0, TRUE ~ event),
          time = case_when(event != 0 ~ (event_td - (d3_td +14))*12/365.25,
                          TRUE ~ (end_dof - (d3_td+14))*12/365.25 ))%>%
  dplyr::select(PID, AGEC, GROUP_NB, CENTER, DOE, time, event, end_dof, d3d4)%>%
   
  mutate(
    #Truncate time for subjects without events
     time = case_when( #14*12/365.25 : 14 days in unit of month
     (event == 0 & is.na(d3d4) == T & time >= month - 14*12/365.25) ~ (month - 14*12/365.25)-0.000000000001,
     (event == 0 & is.na(d3d4) == T & time < month - 14*12/365.25) ~ as.double(time)-0.000000000001,
     (event == 0 & time >= (d3d4-14*12/365.25) & d3d4 <= (18 + 14*12/365.25)) ~ (d3d4-14*12/365.25)-0.000000000001,
     (event == 0 & time >= (d3d4-14*12/365.25) & d3d4 > (18 + 14*12/365.25)) ~ (month - 14*12/365.25)-0.000000000001,
     (event == 0 & time < (d3d4-14*12/365.25)  & d3d4 <= (18 + 14*12/365.25)) ~ as.double(time)-0.000000000001,
     (event == 0 & time < (d3d4-14*12/365.25)  & d3d4 > (18 + 14*12/365.25)) ~ min(month - 14*12/365.25, time)-0.000000000001,
     TRUE ~ as.double(time)),
     #set censoring time
      end_dof = (as.double(end_dof) - as.double(DOE))*12/365.25,
      follow_up = case_when(
            (is.na(d3d4) == T & end_dof >= month - 14*12/365.25) ~ month - 14*12/365.25,
            (is.na(d3d4) == T & end_dof < month - 14*12/365.25) ~ end_dof,
            (end_dof >= (d3d4-14*12/365.25) & d3d4 <= (month + 14*12/365.25)) ~ d3d4-14*12/365.25,
            (end_dof >= (d3d4-14*12/365.25) & d3d4 > (month + 14*12/365.25)) ~ month - 14*12/365.25,
            (end_dof < (d3d4-14*12/365.25) & d3d4 <= (month + 14*12/365.25)) ~ end_dof,
            (end_dof < (d3d4-14*12/365.25) & d3d4 > (month + 14*12/365.25)) ~ min(month - 14*12/365.25, end_dof)),
          GROUP_NB = case_when(GROUP_NB == 3 ~ 0,
                               TRUE ~ 1))%>%
   group_by(PID)%>%
   # Assign start time and end time for multiple events
   mutate(start = lag(time))%>%
   ungroup()%>%
   mutate(start = case_when(is.na(start) == T ~ 0,
                           TRUE ~ start),
          end = case_when(start == 0 & time >= follow_up ~ follow_up -0.000000000001,
                          TRUE ~ time),
          event = case_when(start == 0 & time > follow_up ~ 0,
                            TRUE ~ event))%>%
  filter(end <= follow_up)

# Add a row for non-censored subjects at 20th month (subjects with events)
event_follow <- filter(data.join, event ==1 & abs(end_dof-time) <= 0.00001)
events <- filter(data.join, event ==1 & end_dof > follow_up & PID %in% event_follow$PID == F)%>%
  group_by(PID)%>%
  summarise(AGEC = last(AGEC),
            GROUP_NB = last(GROUP_NB),
            CENTER = last(CENTER),
            DOE = last(DOE),
            event = max(event),
            follow_up = max(follow_up),
            end_dof = max(end_dof))%>%
  mutate(time = follow_up,
         event = 0)
# Add a row for subjects censored before 20th month (subjects with events)
events.c <- filter(data.join, event ==1 & end_dof <= follow_up & PID %in% event_follow$PID == F)%>%
  group_by(PID)%>%
  summarise(AGEC = last(AGEC),
            GROUP_NB = last(GROUP_NB),
            CENTER = last(CENTER),
            DOE = last(DOE),
            event = max(event),
            follow_up = max(follow_up),
            end_dof = max(end_dof),
            time = last(follow_up))%>%
  mutate(event = 0)

analysis.1 <- data.join%>%
   dplyr::select(PID, AGEC, GROUP_NB, CENTER, DOE, event, follow_up, end_dof, time)%>%
   rbind(events, events.c)%>%
   arrange(PID, time)

# Remove events within 14 days of the subject's previous event
data.censor <- filter(analysis.1, event == 0)
data.events <- filter(analysis.1, event == 1)%>%
  group_by(PID)%>%
  mutate(group = accumulate(time, ~ if (.y - .x <= 14*12/365.25+ 0.000000000001) .x else .y))%>%
  ungroup()%>%
  group_by(PID, group)%>%
  slice_head(n = 1)%>%
  ungroup()%>%
  dplyr::select(-group)

analysis_final <- rbind(data.censor, data.events)%>%
  arrange(PID, time)%>%
   group_by(PID)%>%
   # Assign start time and end time for multiple events
   mutate(start = lag(time) + 14*12/365.25)%>%
   ungroup()%>%
   mutate(start = case_when(is.na(start) == T ~ 0,
                           TRUE ~ start),
          end = case_when(start == 0 & time >= follow_up ~ follow_up,
                          TRUE ~ time))%>%
  filter(end - start >0)
  
return(analysis_final)
}

# Follow-up 17.5 months
analysis_longer <- function.finaldata(18)

# Follow-up 14.5 months
analysis_15 <- function.finaldata(15)

# Follow-up 11.5 months
analysis_shorter <- function.finaldata(12)

```

```{r}
# population information of 17.5 month data
nrow(analysis_longer%>%distinct(PID))
nrow(analysis_longer%>%filter(GROUP_NB == 1)%>%distinct(PID))
nrow(analysis_longer%>%filter(GROUP_NB == 0)%>%distinct(PID))
```
```{r}
data.all <- analysis_longer%>%distinct(PID, .keep_all = T)
nrow(filter(data.all, end_dof > 0 & end_dof <= 11.5))
nrow(filter(data.all, end_dof > 0 & end_dof <= 11.5))/nrow(data.all)*100
nrow(filter(data.all, end_dof > 11.5 & end_dof <= 14.5))
nrow(filter(data.all, end_dof > 11.5 & end_dof <= 14.5))/nrow(data.all)*100
nrow(filter(data.all, end_dof > 14.5 & end_dof <= 17.5))
nrow(filter(data.all, end_dof > 14.5 & end_dof <= 17.5))/nrow(data.all)*100
nrow(filter(data.all, end_dof > 17.5))
nrow(filter(data.all, end_dof > 17.5))/nrow(data.all)*100
nrow(data.all)
```

```{r}
data.all.i <- analysis_longer%>%distinct(PID, .keep_all = T)%>%filter(GROUP_NB == 1)
nrow(filter(data.all.i, end_dof > 0 & end_dof <= 11.5))
nrow(filter(data.all.i, end_dof > 0 & end_dof <= 11.5))/nrow(data.all.i)*100
nrow(filter(data.all.i, end_dof > 11.5 & end_dof <= 14.5))
nrow(filter(data.all.i, end_dof > 11.5 & end_dof <= 14.5))/nrow(data.all.i)*100
nrow(filter(data.all.i, end_dof > 14.5 & end_dof <= 17.5))
nrow(filter(data.all.i, end_dof > 14.5 & end_dof <= 17.5))/nrow(data.all.i)*100
nrow(filter(data.all.i, end_dof > 17.5))
nrow(filter(data.all.i, end_dof > 17.5))/nrow(data.all.i)*100
nrow(data.all.i)
```

```{r}
data.all.c <- analysis_longer%>%distinct(PID, .keep_all = T)%>%filter(GROUP_NB == 0)
nrow(filter(data.all.c, end_dof > 0 & end_dof <= 11.5))
nrow(filter(data.all.c, end_dof > 0 & end_dof <= 11.5))/nrow(data.all.c)*100
nrow(filter(data.all.c, end_dof > 11.5 & end_dof <= 14.5))
nrow(filter(data.all.c, end_dof > 11.5 & end_dof <= 14.5))/nrow(data.all.c)*100
nrow(filter(data.all.c, end_dof > 14.5 & end_dof <= 17.5))
nrow(filter(data.all.c, end_dof > 14.5 & end_dof <= 17.5))/nrow(data.all.c)*100
nrow(filter(data.all.c, end_dof > 17.5))
nrow(filter(data.all.c, end_dof > 17.5))/nrow(data.all.c)*100
nrow(data.all.c)
```



Modeling

17.5 months

1. PH
```{r}
Y = Surv(analysis_longer$start, analysis_longer$end, analysis_longer$event == 1)
fit.ph <- coxph(Y ~ GROUP_NB + strata(as.factor(CENTER)) + cluster(PID), data = analysis_longer, control = coxph.control(timefix = F), robust = T)
summary(fit.ph)

1-exp(fit.ph[[1]][[1]])
1-exp(fit.ph[[1]][[1]]-1.96*0.03479)
1-exp(fit.ph[[1]][[1]]+1.96*0.03479)
```


2. time-varing intervention effect and AUC
(i) g(t) = beta0 + beta1*t^0.5
```{r}
Y = Surv(analysis_longer$start, analysis_longer$end, analysis_longer$event == 1)
fit.extend <- coxph(Y ~ GROUP_NB + strata(as.factor(CENTER))  + tt(GROUP_NB) + cluster(PID), data = analysis_longer,
                   tt = function(x, t, ...) x*sqrt(t), control = coxph.control(timefix = F))
```
```{r}
summary(fit.extend)
BIC(fit.extend)
```
```{r}
beta0 <- fit.extend[[1]][[1]]
beta0
beta1 <- fit.extend[[1]][[2]]
beta1

auc.fun <- function(t) 1-exp(beta0+beta1*sqrt(t))

integrate(auc.fun, 0, 11.54)
6.082523/11.54

integrate(auc.fun, 0, 14.54)
6.88184/14.54

integrate(auc.fun, 0, 17.54)
7.351755/17.54
```

(ii) g(t) = beta0+ beta1*log(t)
```{r}
Y = Surv(analysis_longer$start, analysis_longer$end, analysis_longer$event == 1)
fit.extend <- coxph(Y ~ GROUP_NB + strata(as.factor(CENTER))  + tt(GROUP_NB) + cluster(PID), data = analysis_longer,
                   tt = function(x, t, ...) x*log(t), control = coxph.control(timefix = F))
#cbind(zt = x*t, zt2 = x*t^2)
```
```{r}
summary(fit.extend)
BIC(fit.extend)
```
```{r}
beta0 <- fit.extend[[1]][[1]]
beta0
beta1 <- fit.extend[[1]][[2]]
beta1

auc.fun <- function(t) 1-exp(beta0+beta1*log(t))

integrate(auc.fun, 0, 11.54)
5.930797/11.54

integrate(auc.fun, 0, 14.54)
6.802859/14.54

integrate(auc.fun, 0, 17.54)
7.494676/17.54
```




14.5 months

1. PH
```{r}
Y = Surv(analysis_15$start, analysis_15$end, analysis_15$event == 1)
fit.ph <- coxph(Y ~ GROUP_NB + strata(as.factor(CENTER)) + cluster(PID), data = analysis_15, control = coxph.control(timefix = F), robust = T)
summary(fit.ph)

1-exp(fit.ph[[1]][[1]])
1-exp(fit.ph[[1]][[1]]-1.96*0.03649)
1-exp(fit.ph[[1]][[1]]+1.96*0.03649)
```


2. time-varing intervention effect and AUC
(i) g(t) = beta0 + beta1*t^0.5
```{r}
Y = Surv(analysis_15$start, analysis_15$end, analysis_15$event == 1)
fit.extend <- coxph(Y ~ GROUP_NB + strata(as.factor(CENTER))  + tt(GROUP_NB) + cluster(PID), data = analysis_15,
                   tt = function(x, t, ...) x*sqrt(t), control = coxph.control(timefix = F))
#cbind(zt = x*t, zt2 = x*t^2)
```
```{r}
summary(fit.extend)
BIC(fit.extend)
```
```{r}
beta0 <- fit.extend[[1]][[1]]
beta0
beta1 <- fit.extend[[1]][[2]]
beta1

auc.fun <- function(t) 1-exp(beta0+beta1*sqrt(t))

integrate(auc.fun, 0, 11.54)
6.020478/11.54

integrate(auc.fun, 0, 14.54)
6.631161/14.54

integrate(auc.fun, 0, 17.54)
6.81528/17.54
```


(i) g(t) = beta0 + beta1*log(t)
```{r}
Y = Surv(analysis_15$start, analysis_15$end, analysis_15$event == 1)
fit.extend <- coxph(Y ~ GROUP_NB + strata(as.factor(CENTER))  + tt(GROUP_NB) + cluster(PID), data = analysis_15,
                   tt = function(x, t, ...) x*log(t), control = coxph.control(timefix = F))
#cbind(zt = x*t, zt2 = x*t^2)
```
```{r}
summary(fit.extend)
BIC(fit.extend)
```
```{r}
beta0 <- fit.extend[[1]][[1]]
beta0
beta1 <- fit.extend[[1]][[2]]
beta1

auc.fun <- function(t) 1-exp(beta0+beta1*log(t))

integrate(auc.fun, 0, 11.54)
5.918475/11.54

integrate(auc.fun, 0, 14.54)
6.769329/14.54

integrate(auc.fun, 0, 17.54)
7.433686/17.54
```




11.5 months

1. PH
```{r}
Y = Surv(analysis_shorter$start, analysis_shorter$end, analysis_shorter$event == 1)
fit.ph <- coxph(Y ~ GROUP_NB + strata(as.factor(CENTER)) + cluster(PID), data = analysis_shorter, control = coxph.control(timefix = F), robust = T)
summary(fit.ph)

1-exp(fit.ph[[1]][[1]])
1-exp(fit.ph[[1]][[1]]+1.96*0.03736)
1-exp(fit.ph[[1]][[1]]-1.96*0.03736)
```


2. time-varing intervention effect and AUC
(i) g(t) = beta0 + beta1*t^0.5
```{r}
Y = Surv(analysis_shorter$start, analysis_shorter$end, analysis_shorter$event == 1)
fit.extend.short <- coxph(Y ~ GROUP_NB + strata(as.factor(CENTER)) + tt(GROUP_NB) + cluster(PID), data = analysis_shorter, tt = function(x, t, ...) x*sqrt(t),
                          control = coxph.control(timefix = F))
#cbind(zt = x*t, zt2 = x*t^2)
```
```{r}
summary(fit.extend.short)
BIC(fit.extend.short)
```
```{r}
beta0 <- fit.extend.short[[1]][[1]]
beta0
beta1 <- fit.extend.short[[1]][[2]]
beta1

auc.fun <- function(t) 1-exp(beta0+beta1*sqrt(t))

integrate(auc.fun, 0, 11.54)
5.901259/11.54

integrate(auc.fun, 0, 14.54)
6.309475/14.54

integrate(auc.fun, 0, 17.54)
6.188608/17.54
```

(ii) g(t) = beta0 + beta1*log(t)
```{r}
Y = Surv(analysis_shorter$start, analysis_shorter$end, analysis_shorter$event == 1)
fit.extend.short <- coxph(Y ~ GROUP_NB + strata(as.factor(CENTER)) + tt(GROUP_NB) + cluster(PID), data = analysis_shorter, tt = function(x, t, ...) x*log(t),
                          control = coxph.control(timefix = F))
#cbind(zt = x*t, zt2 = x*t^2)
```
```{r}
summary(fit.extend.short)
BIC(fit.extend.short)
```
```{r}
beta0 <- fit.extend.short[[1]][[1]]
beta0
beta1 <- fit.extend.short[[1]][[2]]
beta1

auc.fun <- function(t) 1-exp(beta0+beta1*log(t))

integrate(auc.fun, 0, 11.54)
5.95057/11.54

integrate(auc.fun, 0, 14.54)
6.828754/14.54

integrate(auc.fun, 0, 17.54)
7.526842/17.54
```






































